Timescales for the Effects of Interactions on Galaxy Properties and SMBH Growth

### 2024-06-14 paper-source/main.tex
Major updates to code. Need to update paper as well. Thesis subitted with everything looking good.

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Interactions in COSMOS]{When do the effects of interaction happen? Interacting and disturbed galaxies in the COSMOS field}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[D. O'Ryan et al.]{
David O'Ryan$^{1}$\thanks{E-mail: d.oryan@lancaster.ac.uk}
\\
% List of institutions
$^{1}$Lancaster University \\
}

@@

% Abstract of the paper
\begin{abstract}
   Galaxy interaction and merging have fundamental impacts on galactic evolution. From increases in star formation, to morphology, to activation of nuclear activity, galaxy interaction is thought to play a role. However, at what point in the interaction do these effects occur? In this work, we divide a large sample of interacting galaxies in the COSMOS field into four distinct stages based on their interaction morphology. Each stage captures a different part of the dynamical timescale of interaction. We then probe the relation between stage, mass, star formation, and nuclear activity. We find that the distribution of star formation rates rapidly changes with stage for an equivalent mass distribution. We find a shrinking red sequence with stage, and an increase in the fraction of starburst galaxies from 0.1 as close pairs to 0.38 at the point of coalescence. We also find that the fraction of active galactic nuclei is static through most of the interaction, until it increases at coalescence. We discuss these results, and compare them with the more prevalent stage diagnostic of projected separation between galaxy pairs. We find morphology and stage dependence in the rate of star formation which is not captured by using the projected separation. We show that the point in time of the underlying dynamical timescale an interaction occupies is as important to consider as its projected separation.

\end{abstract}

@@

\section{Introduction}\label{introduction}
\noindent To fully map out the effect of galaxy interaction, we must understand its impact through various merger histories and across the full dynamical timescale of the interaction. The dynamical timescale is the full time of the interaction, from the two galaxies approaching each other to flying by and escaping or merging. We cannot observe a full, continuous interaction and have to rely on many observations to provide snapshots of the dynamical timescale we can piece together. This, however, makes it difficult to conclusively find link the dynamical timescale to physical processes. Often, observational samples are too small to be fully representative or deep studies of individual systems are required to identify the dynamical timescale \citep{2000ApJ...530..660B, 2012A&A...539A..45L, 2014A&A...567A.132W}. We gain insights from works focused on simulations, where the dynamical timescale can be directly recorded. They find sudden increases in star formation occurs at the mid-point of the dynamical timescale, as the two systems flyby \citep{2008MNRAS.384..386C, 2019MNRAS.490.2139R, 2021MNRAS.503.3113M}. However, confirming this observationally remains elusive \citep{2023ApJ...958...96R}. Observational confirmation is also difficult with nuclear activation and quenching of interacting systems \citep{2011MNRAS.418.2043E, 2018PASJ...70S..37G, 2023ApJ...942..107S}.

While cannot directly measure the full dynamical timescale for individual systems, we can create large samples of interacting galaxies contain systems spanning the full dynamical history. Attempts to create such a sample have been either by machine learning classification \citep{2019A&A...626A..49P, 2023A&A...669A.141S}, visual classification by citizen scientists \citep{2010MNRAS.401.1043D} or by photometric parameterisation \citep{2004AJ....128..163L, 2023MNRAS.522....1N}. However, as stated previously, each sample has always been plagued by contamination and a loss of statistical significance when broken down into stages.

@@

In this Chapter, we split our large sample of interacting galaxies found in Chapter \ref{chapter2} into four specific stages based on their morphology. Each stage is designed to capture different parts of the dynamical time of an interaction. The stages range from galaxies being close pairs, to overlapping and disturbed, to disturbed and distinct to finally coalescing. This follows the classification methods of developing algorithms for staging of interaction\citep{2019MNRAS.490.5390B,2022ApJ...937...97C}.

@@

This Chapter is laid out as follows in Chapter \ref{data} we briefly summarise the COSMOS catalogue, and describe the process of catalogue matching especially focused on accurate de-duplication and cross-matching. Chapter \ref{method} describes the methods by which we split the interacting systems into stages, and how we identify the secondary galaxies in each pair. In Sections \ref{results:SF_stage} and \ref{results:AGN_stage} we show our results of star formation and active galactic nuclei evolution with interaction stage and present an initial discussion. These are followed in Chapter \ref{discussion} where we compare to previous works and put our results in the context of the field. Finally, in Chapter \ref{conclusion} we make concluding remarks and discuss future work to better our constraints.

\section{Data: Catalogue Matching \& Secondary Identification} \label{data}
\subsection{The COSMOS2020 Catalogue}
In the COSMOS survey, there are multiple catalogues of galactic parameters such as star formation rates (SFR), stellar masses and line emissions. We elect to specifically use the COSMOS2020 catalogue \citep{2022ApJS..258...11W}. This catalogue has a wealth of ancillary information approximately 1.7 million sources in a 2 square degree area of the sky. Each source has been analysed with well known astronomical software (for our purposes, primarily \texttt{\texttt{LePhare}} \citep{1999MNRAS.310..540A, 2006A&A...457..841I}, \texttt{\texttt{EAzY}} \citep{2008ApJ...686.1503B} and \texttt{\texttt{FAST}} \citep{2017MNRAS.465.3390A}). These provide estimates of the physical parameters of each source based on its measured broadband photometry. For our purposes, the parameters of interest are the stellar masses and the estimated star formation rates of each source. For the stellar mass we use the best fit \texttt{LePhare} measurement and for the star formation rates we use the best fit \texttt{EAzY}/\texttt{FAST} measurements. We compare the output masses and SFRs for our sources in Figure \ref{fig:difference-measures}. As shown, there can be wide scatter and disagreement between the two algorithms. We tested different combinations of \texttt{EAzY} and \texttt{LePhare} stellar masses and SFRs to find which would be best to use in this work. We found that using the same algorithm in both parameters revealed the degeneracies of using discrete templates to fit the broad band photometry in the underlying parameter space. Therefore, we elect to inter-mix the algorithms and use the best fit \texttt{LePhare} stellar mass and best fit \texttt{EAzY} SFR measurement.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/mass-sfr-scatter.pdf}
\caption[Comparison of the measures of stellar mass and SFR using either \texttt{LePhare} or \texttt{EAzY}/\texttt{FAST} photometric codes to calculate them.]{Comparison of the measures of stellar mass and SFR using either \texttt{LePhare} or \texttt{EAzY}/\texttt{FAST} photometric codes to calculate them. If the algorithms agreed perfectly, the sources would lie on the blue 1:1 line. \textit{Left}: The scatter in the stellar masses between softwares. As shown, \texttt{EAzY} often seems to find larger stellar masses when compared to \texttt{LePhare}. \textit{Right}: Scatter in SFRs when measured with \texttt{LePhare} or \texttt{EAzY}.}
\label{fig:difference-measures}
\end{figure}

@@

Once matched, we remove any sources with non-physical photometric measurements for the stellar mass or star formation rates. We then further reduce our sample by only keeping sources within a mass range of $6.5 \leq \log_{10} \text{M}_{*}(\text{M}_{\odot}) \leq 12.5$ and a star formation rate range of $-5 < \log_{10} \text{SFR} (\text{M}_{\odot}\text{yr}^{-1}) \leq 3.5$. We also opt to institute a redshift cut of $z \leq 1.2$. Beyond this redshift, we find that identification of tidal features becomes difficult due to surface brightness dimming and we risk mis-identifying the stages of the interacting galaxies. This also matches the redshift cut applied to the environment catalogue we cross match with in Chapter \ref{data:environ}. Applying these cuts reduces our sample size to 3,689 interacting galaxies.

\subsection{Secondary Identification}\label{sec:sec-ident}
\noindent As the catalogue described in Chapter \ref{chapter2} only contains source coordinates and IDs, we must also manually identify the secondaries of many of our interacting systems. To find the secondaries, we apply three steps. First, a cutout surrounding each source was created. These cutouts were from the COSMOS cutout service, selecting HST-ACS tiles in the $F814W$ filter. Each cutout had a 30" radius (corresponding to 1001 $\times$ 1001 pixels). The original cutout from Chapter \ref{chapter2} was also displayed next to the enlarged cutout. We annotate each cutout with each sources COSMOS2020 ID and measured photometric redshift and error. By annotating each cutout with the sources photometric redshift and ID, we visually assessed each cutout and gave one of the four following classifications to each: system disturbed but secondary could not be identified; secondary could be identified; cannot confirm galaxy is interacting; null redshift (0 or NaN); incorrect primary assigned.

@@

% Define our classifications above.
While initially surprising that the majority of our systems could not have a secondary identified, we found that it was mostly due to limitations in the COSMOS catalogue or the way in which we conduct our secondary identification. Each potential secondary must have a COSMOS ID associated with it, however, when two systems are very close together and small enough they were identified under a single COSMOS ID despite being two separate systems. The same also occurred when two systems were merging or interacting. The tidal features connecting the systems or coalescing systems would only be identified under one ID in the catalogue. Figure \ref{fig:secondary_selection} panel C) shows an example of two systems being close enough together that they have been identified under a single COSMOS2020 entry. Figure \ref{fig:secondaries_found} shows this disparity with the different types of interaction we observe.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/visualisation_classification.pdf}
\caption[Where a secondary could be identified at different stages in the interaction.]{Where a secondary could be identified at different stages in the interaction. The reasoning for such disparity in secondaries identified is due to the relative distance each the secondary would be from the primary at each stage. For a close pair, we often found the secondary galaxy, but a minority of these were so close together that the entire system was given a single COSMOS ID. The same was true for those interacting systems approaching pericentre or merging. When the secondary was near apocentre, often it would be outside the cutout we were using for visual classification.}
\label{fig:secondaries_found}
\end{figure}

@@

For each galaxy in our sample, we define a mass- and redshift-matched control galaxy to investigate differences in their galactic parameters. We find these control galaxies from the COSMOS2020 catalogue. All galaxies within 0.01 dex of our samples stellar mass are selected from the catalogue, and within a $\pm0.01$ redshift slice. We then define the control galaxy as the system furthest from our interacting galaxy. Each control galaxy is then visually confirmed to be non-interacting and have no nearby pairs. We also confirm that it does not already exist in the catalogue, and ensure there is no duplication in the control sample. Figure \ref{fig:matched-distributions} shows the mass distribution of our paired sub-sample of galaxies and their control, showing a mass distribution which is approximately the same as the interacting sample. With the previously defined cuts, we find a control galaxy for all but one of our interacting galaxies.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/mass-matching-pairs.pdf}
\caption[The mass distribution of the paired interacting galaxy sample and the control sample.]{The mass distribution of the paired interacting galaxy sample and the control sample. Both primary and secondary galaxies are within this distribution.}
\label{fig:matched-distributions}
\end{figure}

\subsection{Finding Additional Systems}
\noindent As a result of using visual classification to find the secondary galaxy in each interaction, we were able to also confirm other interacting systems which had not been found in our catalogue. Primarily, these extra interacting galaxies are from systems which had more than two galaxies involved in the interaction. Our selection process was built to only find a primary and secondary galaxy and, therefore, we add these extra systems into our sample manually. Other interacting galaxies that were added were low redshift systems which would have appeared to completely fill the cutout of the classification process in Chapter \ref{chapter2}. By looking at the larger COSMOS2020 cutouts, we are able to recover these galaxies and add them to this sample.

@@

\subsection{Creating the Mass-Limited Sample}
\noindent We now investigate the distribution of our sample between mass and redshift. Figure \ref{fig:redshift_selection} shows the resultant distribution from our flux limited sample. In order to ensure that our results are not biased by a redshift-dependent mass/luminosity limit, we institute a cut in the mass of the interacting systems. Thus, creating a mass limited sample. We elect to use a mass cut of $\log(\frac{M}{M_\odot}) \geq 9.25$, shown by the blue dashed line in Figure \ref{fig:redshift_selection}. This specific cut is motivated by two reasons. First, this is the lowest mass system which is still observed in our the coalescing sample at $z = 1.2$. The second reason is that such a mass cut is approximately the one made in the environment catalogue we will describe later in Chapter \ref{data:environ}. There we make a cut of $\log(\frac{M}{M_\odot}) \geq 9.6$, we find that decreasing the mass cut to $\log(\frac{M}{M_\odot}) \geq 9.25$ does not make major differences to our results. Thus, we opt to use a lower mass cut to increase the number of galaxies in our sample. Making such a cut reduces our flux-limited sample of 4,181 to a mass-limited sample of 3,384 interacting galaxies.

@@

Throughout this Chapter, we will use our mass-limited sample in our analysis. This gives us uniform sensitivity across our volume. However, we also conducted the same analysis described below with the flux-limited sample and the qualitative conclusions of this Chapter did not change.

\subsection{Sample Count Summary}\label{sec:sample-summary}
\noindent Thus, to briefly sumarise this chapter, we have applied multiple cuts and additions to our initial sample of interacting galaxies. We have gone from an initial sample of 3,689 matched galaxies with COSMOS to a mass-limited sample of 3,384 galaxies. This mass-limited sample of galaxies is the one we will use throughout this work. In the mass-limited paired sample there are 338 pairs or 676 galaxies.

@@

\subsection{Classifying Stage of Interaction}\label{sec:staging}
The primary goal of this work is to find if a relation exists between a host of galactic parameters, underlying physical processes and the stage of the interaction. Each stage covers a different part of the dynamical history in an interaction and, in this case, we define it based on the morphology and projected separation between systems. We have already encountered the four stages we will investigate in Figures \ref{fig:matched-distributions} and \ref{fig:redshift_selection} and now define a short hand name to refer to them as well as fully describe the part of the dynamical history they represent. These are:

@@

Figure \ref{fig:stages} shows the original source cutouts used in Chapter \ref{chapter2} to give an example of each stage. There are degeneracy's associated with this staging system, however. In this context, we define a degeneracy as when the interacting galaxies may be at two or more parts of the dynamical timescale and we have no way to tell without further information on the system.

@@

We have noted the degeneracy of each stage as we have described them. Thus, we can now put this fully into the context of the dynamical timescale. Over a typical interaction, we would expect the galaxy pair to move from being separated to the pericentre of the interaction in the early times of the dynamical timescale. Then, dependent on the velocity of the system, the galaxy pair will either move straight into the merging stage of classification and begin to coalesce or it will move towards the apocentre stage. This change from pericentre to apocentre can then take two branching paths dependent on the velocity in the system. If the galaxies have sufficient velocity, the apocentre stage will be their end state until the tidal features slowly dissipate. If they do not have enough velocity to escape on another, the system will move from the apocentre stage back into the pericentre stage. This could then happen for many cycles until finally the galaxies enter the merging stage and coalesce. Figure \ref{fig:illustration} shows the branching paths that the galaxies can take through each stage. These images are created using the Advanced Python Stellar Animation Module restricted numerical simulation, described in Chapter \ref{chapter4}, and are for illustrative purposes only here.

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figures/stage-evolution.jpg}
\caption[The progression through an interaction using our stage definitions.]{The progression through an interaction using our stage definitions. In the separated stage, we have two systems that are approaching each other but exhibit no tidal features. This is before the point of closest approach has occurred. This is followed by an intial pericentre stage: the systems are approximately at their closest approach. This stage is often mistaken, in the COSMOS catalogue, for being of only one source. Clear tidal features exist with major disturbance in the two disks. This is followed by the apocentre stage, where there are two distinct cores with clear tidal features. However, after this point, there are two outcomes to the system depending on the galactic velocities. If the secondary has the escape velocity, the system will remain an apocentre stage interaction until the tidal features dissipate (and no longer are in our sample). If they do not have enough energy, the system will return to the pericentre stage of the encounter and then begin to coalesce in the merging stage. Images are from the Advanced Python Stellar Particle Animation Module interacting galaxy algorithm described in Chapter \ref{chapter4} and based on the stellar particle animation module algorithm described in \citet{2016A&C....16...26W}.}
\label{fig:illustration}
\end{figure*}

@@

% Need to integrate this figure.
\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figures/redshift-proj-sep-diagnostic.pdf}
\caption[The measured distribution of projected separation with redshift.]{The measured distribution of projected separation with redshift. As we have made redshift cuts out to $z = 1.2$, we investigate the limitations on the projected separations we can successfully identify across our volume. We find that we can successfully identify secondary galaxies and their projected separations down to 10kpc to $z=1.2$. This is true across defined interaction stage. We find, because of the defined size of our cutouts, the limitation of finding the secondary galaxy is at low redshift.}
\label{fig:proj-seps-limits}
\end{figure*}

@@

There is no measure of the environmental density in the COSMOS2020 catalogue. Such a measure is often calculated in numerous ways, such as the N-nearest neighbour \citep{2006MNRAS.373..469B}, different Bayesian metrics \citep{2008ApJ...674L..13C} or estimating it from Voronoi Tesselation \citep{2021inas.book...57V}. However, in this Chapter, we use the existing environmental density catalogue produced by \citet{2017ApJ...837...16D}. This catalogue was created specifically for the COSMOS survey, and has a measured density for all sources with mass $\log(\frac{M}{M_\odot}) \geq 9.6$ and $z \leq 1.2$. \citet{2017ApJ...837...16D} calculate not only the density, but also the density parameter $\delta$ and assign each source to a field, filament or cluster classification. For a full description of how they calculate the environment and density field see \citet{2015ApJ...805..121D} and \citet{2017ApJ...837...16D}, but we will briefly describe it here.

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/redshift_scatter.pdf}
\caption[The scatter in photometric redshifts of our paired sample.]{The scatter in photometric redshifts of our paired sample. Left shows the scatter in the primary and secondary photometric redshift measurements. Right the measured difference in the primary and secondary photometric redshifts found in the left panel. The red dashed lines show the expected error at the maximum redshift of our sample. As shown, the bulk of the sample lies within this region.}
\label{fig:redshift-scatter}
\end{figure}

>>>>

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Timescales of Interaction]{Timescales for the Effects of Interactions on Galaxy Properties and SMBH Growth}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[D. O'Ryan et al.]{David O'Ryan$^{1}$\thanks{E-mail: \href{mailto:d.oryan@lancaster.ac.uk}{d.oryan@lancaster.ac.uk}},
    Brooke Simmons$^{1}$,
    COSMOS Team,
    GZ Team, 
    \newauthor
    Others from OR+23,
\\
% List of institutions
$^{1}$Lancaster University \\
}

@@

% Abstract of the paper
\begin{abstract}
   Galaxy interactions and mergers fundamentally impact galaxy evolution in multiple ways. At what point in the interaction do these effects occur? In this work, we divide a large sample of interacting galaxies observed by the Hubble Space Telescope into four distinct stages based on their interaction morphology. Each stage captures a different part of the dynamical timescale of interaction. We select a sample of 4,181 interacting and merging systems in the COSMOS field, using the rich ancillary data available in this field to probe the relation between merger stage, stellar mass, star formation rate, and AGN activity. We find that the distribution of star formation rates rapidly changes with stage for an equivalent mass distribution. We find a shrinking red sequence with stage, and an increase in the fraction of starburst galaxies from 10\% as close pairs to 32\% at the point of coalescence. We also find that the fraction of active galactic nuclei is static through most of the interaction, until it increases at coalescence. We compare these results with the more prevalent stage diagnostic of projected separation between galaxy pairs. We find morphology and stage dependence in the rate of star formation which is not captured by using the projected separation. We show that the point in time of the underlying dynamical timescale an interaction occupies is as important to consider as its projected separation.
\end{abstract}

@@

\section{Introduction}\label{introduction}
\noindent Galaxy interaction and merging is of fundamental importance to our current theories of galaxy formation and evolution. Our current cosmological theory, $\Lambda$CDM, dictates that structure is expected to form and assemble hierarchically \citep{1978MNRAS.183..341W, 2001MNRAS.328..726S, 2013MNRAS.436.1765M}. Thus, understanding the underlying processes that occur in galaxy interaction and merging is paramount to understanding cosmological theories. It is well known that many underlying processes are triggered or enhanced due to galaxy interaction. The clearest result of interaction is the disruption of galactic morphology, the acceleration of disks to spheroidal systems and the formation of tidal features \citep{1972ApJ...178..623T, 1977ApJ...212..616T, 2005MNRAS.357..753G, 2009MNRAS.397..802H}, increases in star formation rates when compared to the general galaxy population \citep{1991ApJ...370L..65B, 2006ApJ...652...56B, 2014MNRAS.437.2137S, 2015ApJ...807L..16K} and, potentially, their rapid quenching \citep{2013MNRAS.430.1901H, 2023RAA....23i5026D}. There is also evidence to suggest that interaction plays a role in the ignition of active galactic nuclei (AGN) \citep{2011MNRAS.418.2043E, 2015ApJ...806..219C, 2023MNRAS.523.4164H}. Therefore, with interaction having such an affect on the galaxies we see - over a long time period - understanding galaxy interaction is fundamental to understanding galaxy evolution.

To fully map out the effect of galaxy interaction, we must understand its impact through various merger histories and across the full dynamical timescale of the interaction. The dynamical timescale is the full time of the interaction, from the two galaxies approaching each other to flying by and escaping or merging. We cannot observe a full, continuous interaction and have to rely on many observations to provide snapshots of the dynamical timescale we can piece together. This, however, makes it difficult to conclusively find link the dynamical timescale to physical processes. Often, observational samples are too small to be fully representative or deep studies of individual systems are required to identify the dynamical timescale \citep{2000ApJ...530..660B, 2012A&A...539A..45L, 2014A&A...567A.132W}. We gain insights from works focused on simulations, where the dynamical timescale can be directly recorded. They find sudden increases in star formation occurs at the mid-point of the dynamical timescale, as the two systems flyby \citep{2008MNRAS.384..386C, 2019MNRAS.490.2139R, 2021MNRAS.503.3113M}. However, confirming this observationally remains elusive \citep{2023ApJ...958...96R}. Observational confirmation is also difficult with nuclear activation and quenching of interacting systems \citep{2011MNRAS.418.2043E, 2018PASJ...70S..37G, 2023ApJ...942..107S}.

While we cannot yet directly measure the full dynamical timescale for individual systems, we can create large samples of interacting galaxies contain systems spanning the full dynamical history. Attempts to create such a sample have been either by machine learning classification \citep{2019A&A...626A..49P, 2023A&A...669A.141S}, visual classification by citizen scientists \citep{2010MNRAS.401.1043D} or by photometric parameterisation \citep{2004AJ....128..163L, 2023MNRAS.522....1N}. However, as stated previously, each sample has always been plagued by contamination and a loss of statistical significance when broken down into stages.

@@

In this work, we split our large sample of interacting galaxies found in \citet[][hereafter OR23]{2023ApJ...948...40O} into four specific stages based on their morphology. Each stage is designed to capture different parts of the dynamical time of an interaction. The stages range from galaxies being close pairs, to overlapping and disturbed, to disturbed and distinct to finally coalescing. This follows the classification methods of developing algorithms for staging of interaction\citep{2019MNRAS.490.5390B,2022ApJ...937...97C}.

@@

This paper is laid out as follows in Section \ref{data} we briefly summarise the COSMOS catalogue, and describe the process of catalogue matching especially focused on accurate de-duplication and cross-matching. Section \ref{method} describes the methods by which we split the interacting systems into stages, and how we identify the secondary galaxies in each pair. In Sections \ref{results:SF_stage} and \ref{results:AGN_stage} we show our results of star formation and active galactic nuclei evolution with interaction stage and present an initial discussion. These are followed in Section \ref{discussion} where we compare to previous works and put our results in the context of the field. Finally, in Section \ref{conclusion} we make concluding remarks and discuss future work to better our constraints.

\section{Data: Catalogue Matching \& Secondary Identification} \label{data}
\subsection{The COSMOS2020 Catalogue}
In the COSMOS survey, there are multiple catalogues of galactic parameters such as star formation rates (SFR), stellar masses and line emissions. We elect to specifically use the COSMOS2020 catalogue \citep{2022ApJS..258...11W}. This catalogue has a wealth of ancillary information approximately 1.7 million sources in a 2 square degree area of the sky. Each source has been analysed with well known astronomical software (for our purposes, primarily \texttt{\texttt{LePhare}} \citep{1999MNRAS.310..540A, 2006A&A...457..841I}, \texttt{\texttt{EAzY}} \citep{2008ApJ...686.1503B} and \texttt{\texttt{FAST}} \citep{2017MNRAS.465.3390A}). These provide estimates of the physical parameters of each source based on its measured broadband photometry. For our purposes, the parameters of interest are the stellar masses and the estimated star formation rates of each source. For the stellar mass we use the best fit \texttt{LePhare} measurement and for the star formation rates we use the best fit \texttt{EAzY}/\texttt{FAST} measurements. We found that using the same algorithm in both parameters revealed the degeneracies of using discrete templates to fit the broad band photometry in the underlying parameter space. Therefore, we elect to inter-mix the algorithms and use the best fit \texttt{LePhare} stellar mass and best fit \texttt{EAzY} SFR measurement.

@@

Once matched, we remove any sources with non-physical photometric measurements for the stellar mass or star formation rates. We then further reduce our sample by only keeping sources within a mass range of $6.5 \leq \log_{10} \text{M}_{*}(\text{M}_{\odot}) \leq 12.5$ and a star formation rate range of $-5 < \log_{10} \text{SFR} (\text{M}_{\odot}\text{yr}^{-1}) \leq 3.5$. We also opt to institute a redshift cut of $z \leq 1.2$. Beyond this redshift, we find that identification of tidal features becomes difficult due to surface brightness dimming and we risk mis-identifying the stages of the interacting galaxies. This also matches the redshift cut applied to the environment catalogue we cross match with in Section \ref{data:environ}. Applying these cuts reduces our sample size to 3,689 interacting galaxies.

\subsection{Secondary Identification}\label{sec:sec-ident}
\noindent As the catalogue described in OR23 only contains source coordinates and IDs, we must also manually identify the secondaries of many of our interacting systems. To find the secondaries, we apply three steps. First, a cutout surrounding each source was created. These cutouts were from the COSMOS cutout service, selecting HST-ACS tiles in the $F814W$ filter. Each cutout had a 30" radius (corresponding to 1001 $\times$ 1001 pixels). The original OR23 cutout was also displayed next to the enlarged cutout. We annotate each cutout with each sources COSMOS2020 ID and measured photometric redshift and error. By annotating each cutout with the sources photometric redshift and ID, we visually assessed each cutout and gave one of the four following classifications to each: system disturbed but secondary could not be identified; secondary could be identified; cannot confirm galaxy is interacting; null redshift (0 or NaN); incorrect primary assigned.

@@

% Define our classifications above.
While initially surprising that the majority of our systems could not have a secondary identified, we found that it was mostly due to limitations in the COSMOS catalogue or the way in which we conduct our secondary identification. Each potential secondary must have a COSMOS ID associated with it, however, when two systems are very close together and small enough they were identified under a single COSMOS ID despite being two separate systems. The same also occurred when two systems were merging or interacting. The tidal features connecting the systems or coalescing systems would only be identified under one ID in the catalogue. Figure \ref{fig:secondary_selection}c shows an example of two systems being close enough together that they have been identified under a single COSMOS2020 entry.

% \begin{figure}
    % \centering
    % \includegraphics[width=\columnwidth]{figures/visualisation_classification.pdf}
    % \caption[Where a secondary could be identified at different stages in the interaction.]{Where a secondary could be identified at different stages in the interaction. The reasoning for such disparity in secondaries identified is due to the relative distance each the secondary would be from the primary at each stage. For a close pair, we often found the secondary galaxy, but a minority of these were so close together that the entire system was given a single COSMOS ID. The same was true for those interacting systems approaching pericentre or merging. When the secondary was near apocentre, often it would be outside the cutout we were using for visual classification.}
    % \label{fig:secondaries_found}
% \end{figure}

@@

For each galaxy in our sample, we define a mass- and redshift-matched control galaxy to investigate differences in their galactic parameters. We find these control galaxies from the COSMOS2020 catalogue. All galaxies within 0.01 dex of our samples stellar mass are selected from the catalogue, and within a $\pm0.01$ redshift slice. We then define the control galaxy as the system furthest from our interacting galaxy. Each control galaxy is then visually confirmed to be non-interacting and have no nearby pairs. We also confirm that it does not already exist in the catalogue, and ensure there is no duplication in the control sample. With the previously defined cuts, we find a control galaxy for all but one of our interacting galaxies.

% \begin{figure}
    % \centering
    % \includegraphics[width=\columnwidth]{figures/mass-matching-pairs.pdf}
    % \caption[The mass distribution of the paired interacting galaxy sample and the control sample.]{The mass distribution of the paired interacting galaxy sample and the control sample. Both primary and secondary galaxies are within this distribution.}
    % \label{fig:matched-distributions}
% \end{figure}

\subsection{Finding Additional Systems}
\noindent As a result of using visual classification to find the secondary galaxy in each interaction, we were able to also confirm other interacting systems which had not been found in our catalogue. Primarily, these extra interacting galaxies are from systems which had more than two galaxies involved in the interaction. Our selection process was built to only find a primary and secondary galaxy and, therefore, we add these extra systems into our sample manually. Other interacting galaxies that were added were low redshift systems which would have appeared to completely fill the cutout of the classification process in OR23. By looking at the larger COSMOS2020 cutouts, we are able to recover these galaxies and add them to this sample.

@@

\subsection{Creating the Mass-Limited Sample}
\noindent We now investigate the distribution of our sample between mass and redshift. Figure \ref{fig:redshift_selection} shows the resultant distribution from our flux limited sample. In order to ensure that our results are not biased by a redshift-dependent mass/luminosity limit, we institute a cut in the mass of the interacting systems. Thus, creating a mass limited sample. We elect to use a mass cut of $\log(\frac{M}{M_\odot}) \geq 9.25$, shown by the blue dashed line in Figure \ref{fig:redshift_selection}. This specific cut is motivated by two reasons. First, this is the lowest mass system which is still observed in our the coalescing sample at $z = 1.2$. The second reason is that such a mass cut is approximately the one made in the environment catalogue we will describe later in Section \ref{data:environ}. There we make a cut of $\log(\frac{M}{M_\odot}) \geq 9.6$, we find that decreasing the mass cut to $\log(\frac{M}{M_\odot}) \geq 9.25$ does not make major differences to our results. Thus, we opt to use a lower mass cut to increase the number of galaxies in our sample. Making such a cut reduces our flux-limited sample of 4,181 to a mass-limited sample of 3,384 interacting galaxies.

@@

Throughout this paper, we will use our mass-limited sample in our analysis. This gives us uniform sensitivity across our volume. However, we also conducted the same analysis described below with the flux-limited sample and the qualitative conclusions of this paper did not change.

\subsection{Sample Count Summary}\label{sec:sample-summary}
\noindent Thus, to briefly summarise this Section, we have applied multiple cuts and additions to our initial sample of interacting galaxies. We have gone from an initial sample of 3,689 matched galaxies with COSMOS to a mass-limited sample of 3,384 galaxies. This mass-limited sample of galaxies is the one we will use throughout this work. In the mass-limited paired sample there are 338 pairs or 676 galaxies.

@@

\subsection{Classifying Stage of Interaction}\label{sec:staging}
The primary goal of this work is to find if a relation exists between a host of galactic parameters, underlying physical processes and the stage of the interaction. Each stage covers a different part of the dynamical history in an interaction and, in this case, we define it based on the morphology and projected separation between systems. We have already encountered the four stages we will investigate in Figure \ref{fig:redshift_selection} and now define a short hand name to refer to them as well as fully describe the part of the dynamical history they represent. These are:

@@

Figure \ref{fig:stages} shows the original source cutouts used in OR23 to give an example of each stage. There are degeneracy's associated with this staging system, however. In this context, we define a degeneracy as when the interacting galaxies may be at two or more parts of the dynamical timescale and we have no way to tell without further information on the system.

@@

We have noted the degeneracy of each stage as we have described them. Thus, we can now put this fully into the context of the dynamical timescale. Over a typical interaction, we would expect the galaxy pair to move from being separated to the pericentre of the interaction in the early times of the dynamical timescale. Then, dependent on the velocity of the system, the galaxy pair will either move straight into the merging stage of classification and begin to coalesce or it will move towards the apocentre stage. This change from pericentre to apocentre can then take two branching paths dependent on the velocity in the system. If the galaxies have sufficient velocity, the apocentre stage will be their end state until the tidal features slowly dissipate. If they do not have enough velocity to escape on another, the system will move from the apocentre stage back into the pericentre stage. This could then happen for many cycles until finally the galaxies enter the merging stage and coalesce. Figure \ref{fig:illustration} shows the branching paths that the galaxies can take through each stage. These images are created using the Advanced Python Stellar Animation Module restricted numerical simulation, described in O'Ryan et al., in prep, and are for illustrative purposes only here.

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figures/stage-evolution.jpg}
\caption[The progression through an interaction using our stage definitions.]{The progression through an interaction using our stage definitions. In the separated stage, we have two systems that are approaching each other but exhibit no tidal features. This is before the point of closest approach has occurred. This is followed by an intial pericentre stage: the systems are approximately at their closest approach. This stage is often mistaken, in the COSMOS catalogue, for being of only one source. Clear tidal features exist with major disturbance in the two disks. This is followed by the apocentre stage, where there are two distinct cores with clear tidal features. However, after this point, there are two outcomes to the system depending on the galactic velocities. If the secondary has the escape velocity, the system will remain an apocentre stage interaction until the tidal features dissipate (and no longer are in our sample). If they do not have enough energy, the system will return to the pericentre stage of the encounter and then begin to coalesce in the merging stage. Images are from the Advanced Python Stellar Particle Animation Module interacting galaxy algorithm described in O'Ryan et al., in prep and based on the stellar particle animation module algorithm described in \citet{2016A&C....16...26W}.}
\label{fig:illustration}
\end{figure*}

@@

% Need to integrate this figure.
\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/redshift-proj-sep-diagnostic.pdf}
\caption[The measured distribution of projected separation with redshift.]{The measured distribution of projected separation with redshift. As we have made redshift cuts out to $z = 1.2$, we investigate the limitations on the projected separations we can successfully identify across our volume. We find that we can successfully identify secondary galaxies and their projected separations down to 10kpc to $z=1.2$. This is true across defined interaction stage. We find, because of the defined size of our cutouts, the limitation of finding the secondary galaxy is at low redshift.}
\label{fig:proj-seps-limits}
\end{figure}

@@

There is no measure of the environmental density in the COSMOS2020 catalogue. Such a measure is often calculated in numerous ways, such as the N-nearest neighbour \citep{2006MNRAS.373..469B}, different Bayesian metrics \citep{2008ApJ...674L..13C} or estimating it from Voronoi Tesselation \citep{2021inas.book...57V}. However, in this Section, we use the existing environmental density catalogue produced by \citet{2017ApJ...837...16D}. This catalogue was created specifically for the COSMOS survey, and has a measured density for all sources with mass $\log(\frac{M}{M_\odot}) \geq 9.6$ and $z \leq 1.2$. \citet{2017ApJ...837...16D} calculate not only the density, but also the density parameter $\delta$ and assign each source to a field, filament or cluster classification. For a full description of how they calculate the environment and density field see \citet{2015ApJ...805..121D} and \citet{2017ApJ...837...16D}, but we will briefly describe it here.

@@

% \begin{figure}
% \centering
% \includegraphics[width=\columnwidth]{figures/redshift_scatter.pdf}
% \caption[The scatter in photometric redshifts of our paired sample.]{The scatter in photometric redshifts of our paired sample. Left shows the scatter in the primary and secondary photometric redshift measurements. Right the measured difference in the primary and secondary photometric redshifts found in the left panel. The red dashed lines show the expected error at the maximum redshift of our sample. As shown, the bulk of the sample lies within this region.}
% \label{fig:redshift-scatter}
% \end{figure}
